A star formation event in the leading arm of the Magellanic stream triggered by passage through the outer Galactic disk

### 2018-08-03 ms.tex
fix query

% Shortcuts for this paper
\newcommand{\clustername}{\textsl{NAME}}
\newcommand{\lmcsmc}{LMC--SMC}
\newcommand{\bprp}{\ensuremath{\textrm{BP} - \textrm{RP}}}
\newcommand{\truepm}{\ensuremath{\tilde{\bs{\mu}}}}
\newcommand{\isodist}{\ensuremath{25~\kpc}}

@@

We report the discovery of a young ($\lesssim 100~\textrm{Myr}$), metal-poor ($[\textrm{Fe}/\textrm{H}] \sim -1$) stellar association in the halo of the Milky Way ($(R, z) \sim (24, 16)~\textrm{kpc}$).
The association --- \clustername\ --- is likely associated with the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with HI gas in the stream, we directly measure the distance to the Magellanic stream, $d \approx 25~\textrm{kpc}$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way. \todo{that don't account for ram pressure and gas interaction with MW disk}
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane.
We therefore conclude that this star-formation event was triggered by the last disk passage, which occurred at a Galactocentric radius $R \approx XX~\kpc$.

\end{abstract}

@@

We identify probable members of the photometrically-identified young cluster using astrometric data from \gaia.
We start by selecting all stars with $\bprp < 0.35$ to remove low-mass and old main sequence star contamination in the region.
\figurename~\ref{fig:pm-members}, left (grey points), shows the sky positions of stars that pass this blue cut in the region around the young cluster.
The larger solid-line (green) circle indicates the region we define as the cluster area, and the two smaller dashed-line (blue) circles indicate control fields that are combined and used for modeling the background distribution of proper motions.
The control fields are designed to, together, have the same total area as the cluster field.
The top-left panel of the middle panels shows the proper motion distribution in the control fields, and top-right panel in the middle shows the proper motion distribution of the cluster field.
The distinct over-density of stars near $(\mu_\alpha, \mu_\delta) \approx (-0.5, 0.5)~\masyr$\footnote{Throughout this article, we use $\mu_\alpha$ to refer to the proper motion value provided by \gaia, which includes the $\cos\delta$ term.} is the identified comoving association of blue stars.

@@

Query to retrieve \gaia\ data around the blue, comoving group found and discussed in \sectionname~\ref{sec:data}:
\begin{verbatim}
SELECT *
FROM gaiadr2.gaia_source
WHERE parallax < 1
    AND ra > (179-6) AND ra < (179+6)
    AND dec > (-29-6) AND dec < (-29+6)
\end{verbatim}

>>>>

% Shortcuts for this paper
\newcommand{\clustername}{\textsl{Price-Whelan 1}}
\newcommand{\lmcsmc}{LMC--SMC}
\newcommand{\bprp}{\ensuremath{\textrm{BP} - \textrm{RP}}}
\newcommand{\truepm}{\ensuremath{\tilde{\bs{\mu}}}}
\newcommand{\isodist}{\ensuremath{25~\kpc}}

@@

We report the discovery of a young, $\lesssim 100~\textrm{Myr}$, metal-poor, $[\textrm{Fe}/\textrm{H}] \sim -1$, comoving stellar association in the halo of the Milky Way, $(R, z) \sim (24, 16)~\textrm{kpc}$.
The association --- \clustername\ --- is likely associated with the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
By assuming that the cluster is co-located with HI gas in the stream, we directly measure the distance to the Magellanic stream, $d \approx 25~\textrm{kpc}$.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way. \todo{that don't account for ram pressure and gas interaction with MW disk}
The estimated age of \clustername\ is consistent with the time of last passage through the Galactic midplane.
We therefore conclude that this star-formation event was triggered by the last disk passage, which occurred at a Galactocentric radius $R \approx XX~\kpc$.
\todo{Further follow-up needed...spectroscopy}
\end{abstract}

@@

We select probable members of the photometrically-identified young cluster using astrometric data from \gaia.
We start by selecting all stars with $\bprp_0 < 0.35$ to remove low-mass and old main sequence star contamination in the region.
\figurename~\ref{fig:pm-members}, left (grey points), shows the sky positions of stars that pass this blue cut in the region around the young cluster.
The larger solid-line (green) circle indicates the region we define as the cluster area, and the two smaller dashed-line (blue) circles indicate control fields that are combined and used for modeling the background distribution of proper motions.
The control fields are designed to, together, have the same total area as the cluster field.
The top-left panel of the middle panels shows the proper motion distribution in the control fields, and top-right panel in the middle shows the proper motion distribution of the cluster field.
The distinct over-density of stars near $(\mu_\alpha, \mu_\delta) \approx (-0.5, 0.5)~\masyr$\footnote{Throughout this article, we use $\mu_\alpha$ to refer to the proper motion value provided by \gaia, which includes the $\cos\delta$ term.} is the identified comoving association of blue stars.

@@

Query to retrieve \gaia\ data around the blue, comoving group found and discussed in \sectionname~\ref{sec:data}:
\begin{verbatim}
SELECT *
FROM gaiadr2.gaia_source
WHERE (parallax < 1 OR parallax IS NULL)
    AND ra > 173.5 AND ra < 185.5
    AND dec > -35.0 AND dec < -23.0
\end{verbatim}
